### 2020-05-28 Manuscript.Rmd
Init

---
title: "An Audio-Based 3D Spatial Guidance AR System for Blind Users"
titlerunning: Audio-Based 3D Spatial Guidance
authorrunning: J. Couhlan et al.
output:
  pdf_document:
    template: res/templates/markdown.tex
    toc: no
    keep_tex: yes
institutes:
- id: 1
  name: The Smith-Kettlewell Eye Research Institute, San Francisco, CA
  emails: coughlan@ski.org, brandon.biggs@ski.org, hshen@ski.org
- id: 2
  name: LITIS, University of Rouen, Normandy, France
  emails: marc-aurele.riviere@univ-rouen.fr
keywords:
- Assisitive Devices
- Augmented Reality
- Auditory Substitution
- Visual Impairment
abstract: Augmented reality (AR) has great potential for blind users because it enables
  a range of applications that provide audio information about specific locations
  or directions in the user’s environment. For instance, the CamIO (“Camera Input-Output”)
  AR app makes physical objects (such as documents, maps, devices and 3D models) accessible
  to blind and visually impaired persons by providing real-time audio feedback in
  response to the location on an object that the user is touching (using an inexpensive
  stylus). An important feature needed by blind users of AR apps such as CamIO is
  a 3D spatial guidance feature that provides real-time audio feedback to help the
  user find a desired location on an object. We have devised a simple audio interface
  to provide verbal guidance towards a target of interest in 3D. The experiment we
  report with blind participants using this guidance interface demonstrates the feasibility
  of the approach and its benefit for helping users find locations of interest.
authors:
- name: James Coughlan
  orcid: 0000-1111-2222-3333
  inst: 1
- name: Brandon Biggs
  orcid: 0000-1111-2222-3333
  inst: 1
- name: Marc-Aurèle Rivière
  orcid: 0000-0002-5108-3382
  inst: 2
- name: Huiying Shen
  orcid: 0000-1111-2222-3333
  inst: 1
ressources: res/
templates: res/templates/
bibliography: res/refs.bib
---

```{r setup, include=FALSE}

#TODO: DESCRIPTION with package metadata ?

knitr::opts_chunk$set(
  fig.show="hold",
  fig.align="default",
  warning = FALSE,
  echo = FALSE,
  cache = FALSE,
  error = FALSE
)

colorize <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{%s}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'>%s</span>", color, x)
  } else x
}
```

# State of the Art and Related Technology {#sota}

Many people who are blind or visually impaired have difficulties accessing a range of everyday objects, including printed documents, maps, infographics, appliances and 3D models used in STEM education, needed for daily activities in schools, the home and the workplace. This limits their participation in many cultural, professional and educational activities. While braille labeling is often a useful means of providing access to such objects, there is only limited space for braille, and this method is only accessible to those who can read braille.

Audio labels are a powerful supplement or alternative to braille labels. Standard methods of implementing audio labels require the use of special hardware and materials, which is costly and limits their adoption. For instance, tactile graphics may be overlaid on a touch-sensitive tablet \cite{TalkingTactileTablet}. Documents or other surfaces can be covered with a special form of paper readable by a “smart pen” \cite{mieleTalkingTactileApps2010}. The PenFriend \cite{kendrickPenFriendTouchMemo2011} stylus uses small barcodes affixed to an object to define audio labels.

There is growing interest in using computer vision-based approaches for audio labeling. In these approaches a camera tracks the user’s fingers or handheld pointer as they explore an object, enabling audio labeling for existing objects with minimal or no customization. Past computer vision approaches to audio labeling include \cite{shiMarkitTalkitLowBarrier2017}, which focuses on 3D printed objects, \cite{thevinAugmentedRealityPeople2018}, which uses a depth camera to access flat documents, and \cite{fuscoTactileGraphicsHelper2015}, which provides access to tactile graphics to students in educational settings.

We build on our past work on CamIO \cite{shenCamIO3DComputer2013,coughlanEvaluatingAuthorUser2017,coughlanTowardsAccessibleAudioLabeling2020}, which is short for “Camera Input-Output.” We have implemented CamIO as a stand-alone iPhone app that tracks an inexpensive wooden stylus held by the user to point at locations of interest on an object (called hotspots). When the user points at a hotspot with the stylus tip, relevant audio information, such as text-to-speech or other sounds, is announced.

Experiments with blind participants have demonstrated the feasibility of using CamIO to access existing audio labels. Moreover, our approach allows a visually impaired user to create audio labels independently \cite{coughlanTowardsAccessibleAudioLabeling2020}. However, an ability that CamIO still lacks is the ability to provide real-time guidance to a hotspot. Without such guidance, the user must search the entire surface of an object with the stylus tip until they hear the hotspot they are seeking. This need is particularly acute in cases where the surface is lacking in tactile cues, such as the touch screen buttons on a microwave oven \cite{coughlanTowardsAccessibleAudioLabeling2020}.

Guidance not only saves the user time in exploring large or intricate objects, or objects lacking in tactile features, but can enable entirely new ways of interacting with objects. For instance, a geography learning app could ask a student to point to a specific city on a world globe, and if the student has trouble finding it, the app could guide them to it in an intuitive way. Other examples could include guiding a user to a specific button on an appliance (such as a microwave) to perform a desired function, guiding a student to a specific element on the periodic table, or even guiding a repairperson to find a part of a machine with limited visibility.
Our guidance system is inspired by recent work on auditory and haptic displays for 3D spatial guidance \cite{mayAuditoryDisplaysFacilitate2019,guezou-philippePrototypingEvaluatingSensory2018}, which provide real-time audio/haptic feedback to guide a user towards a target in a virtual environment. We initially developed an interface similar to \cite{mayAuditoryDisplaysFacilitate2019} for CamIO, but after preliminary testing we devised a simpler and more effective interface for our application, which we evaluated experimentally with four blind participants.

# Overview of the CamIO System {#overview}

CamIO is an augmented reality (AR) iPhone app that uses computer vision to provide audio labels for rigid 3D objects. The rigid object of interest is mounted on a board, which is a printed barcode pattern (\ref{fig:exp}) that allows CamIO to estimate the object’s pose (3D translation and 3D orientation) relative to the camera. The iPhone is typically mounted on a tripod or a gooseneck cellphone holder \ref{fig:exp} to obtain a clear view of the board and object, but in some cases \cite{coughlanTowardsAccessibleAudioLabeling2020} users may hold the iPhone by hand. The user points the tip of a passive wooden stylus \ref{fig:exp}, covered with another barcode pattern, to different locations on an object. CamIO estimates the stylus tip’s 3D location relative to the camera and uses the pose estimate of the board to determine the stylus tip’s 3D location relative to the object. Whenever the tip is close enough to a pre-defined hotspot, it triggers audio information about the hotspot. (In other words, hotspot locations are defined relative to the board, which is mounted rigidly to the object.)

A second stylus (not shown, identical in form but distinct from the one shown in \ref{fig:exp}) is used to create new hotspots, and allows the user to make an audio recording to associate with each new hotspot. Our recent work \cite{coughlanTowardsAccessibleAudioLabeling2020} demonstrates the accessibility of this authoring process, which allows blind users to create their own annotations (hotspots and associated audio labels) independently. However, in the study presented in this paper, the experimenter was the only one who created hotspots, since the emphasis here is on providing guidance to pre-existing hotspots.

# 3D Spatial Guidance Interface {#material}

We build on the auditory display described in \cite{mayAuditoryDisplaysFacilitate2019}, which is intended for audio-based 3D targeting of locations in a large virtual workspace (such as a kitchen). Accordingly, the first incarnation of our guidance approach used a repeating monaural sound, whose tempo indicated how far the stylus tip is from the target, and with a pitch (low, medium or high) that indicated whether the tip needed to move up or down (or else was aligned properly in the vertical direction).

Informal tests with two blind participants (one female of age 42, who is also participant P1 in the experiment described in the next section, and one male of age 73) showed that this guidance worked satisfactorily in some cases. However, we found that the feedback was too indirect: the participants didn’t always know how to move the stylus to increase the tempo. As a result, the resulting guidance process was unacceptably slow.
Thus we devised a second, more direct guidance approach: verbal directions telling the user to move the stylus along the cardinal directions (up, down, left, right, forward or back). While the optimal direction to move the stylus tip towards the hotspot might entail a combination of these directions (e.g., simultaneously move left and up), for simplicity we issue directions along only one cardinal direction at a time – the direction that will most improve the alignment to the target.

We define the guidance interface as follows. The board (barcode pattern beneath the truck, see \ref{fig:exp}) defines an xyz coordinate frame, with $+x$ = right, $-x$ = left, $+y$ = forward (away from the user), $-y$ = back (towards the user), $+z$ = up and $-z$ = down. Let the desired target hotspot location be denoted $(x*, y*, z*)$, and the current location of the stylus tip be $(x,y,z)$. Define the error vector $e = (x-x*, y-y*, z-z*)$, and let index $a = argmax_i |e_i|$, i.e., a is the index (1, 2 or 3, corresponding to x, y or z, respectively) of the entry in e with the highest absolute value. In other words, a indicates which coordinate (x, y or z) is most discrepant from the target coordinate location $(x*, y*, z*)$. Then the appropriate direction is issued, e.g., “left” if $a = 1$ and $x > x*$, “right” if $a = 1$ and $x < x*$, “up” if $a = 3$ and $z < z*$, etc. Directions are announced roughly twice per second whenever the stylus is visible. The target hotspot is announced when the stylus gets to within approximately 1 cm from the hotspot. No audio feedback is issued when the stylus is not visible.

# Experiment with Blind Participants {#exp}

We conducted an experiment to assess the feasibility of the spatial guidance feedback and to determine whether the feedback speeds up the target search. The experiment \ref{fig:exp} compared how much time a blind participant needed to find hotspots (points of interest) on a 3D object with spatial guidance vs. without it. The object was a large toy fire truck (27 cm x 12 cm x 14 cm), with K=10 distinct hotspot locations defined in advance. The hotspots were tactilely salient features such as lights, hooks or other features that protruded from the truck. Some hotspots were in recessed locations on the truck, such as the shift stick and steering wheel (\ref{fig:exp}) in the front cabin, or on the side of the truck facing away from the participant; these hotspots were difficult to find and access, whether by fingertip or by stylus. In fact, the difficulty of accessing such locations highlights an advantage of the CamIO stylus, whose tip need not be visible to the camera. This allows the user to explore recessed areas or areas that are otherwise invisible to the camera.

```{r exp, out.width=c("55%","45%"), fig.cap="(left) Blind participant uses CamIO to explore a toy fire truck. The participant holds the CamIO stylus (a barcode pattern printed on a square wooden dowel with a pointy tip) to point to possible hotspots on the truck. The truck is glued to the board, a piece of paper with a separate barcode pattern, which is taped to the table. An iPhone 8 running CamIO is held by a gooseneck cellphone holder to capture a view of the truck, board and stylus. (right) Stylus points to steering wheel in the truck cabin. Though the steering wheel is in a recessed location that is not visible to CamIO, the stylus tip location can still be estimated."}

knitr::include_graphics("res/img/fig1a.jpg")
knitr::include_graphics("res/img/fig1b.jpg")
```

Each hotspot location was searched for under two feedback conditions: G (guidance provided) and NG (no guidance). The participant was given up to 120 sec. to find each hotspot, after which a time-out was declared. In order to minimize the influence of any possible learning effects, which could make it easier to find a hotspot the second time it was searched for, we randomized the sequence of hotspots and alternated between NG and G trials. This resulted in a sequence of 2K = 20 trials, starting with an NG trial, the same sequence for each participant.

We had a total of N=4 blind participants in our experiment (2 male/2 female, ages from 26-42). Participants P1, P3 and P4 were already familiar with the CamIO system (P4 is a co-author of this paper) while P2 was unfamiliar with it. None of the participants was told the study design before their trials, and we didn’t tell them whether or not any locations would be repeated in the 20 trials. After obtaining IRB consent, we reviewed how CamIO works if necessary and demonstrated the audio guidance feature. The experimenter launched CamIO on an iPhone 8 and positioned it on a gooseneck cellphone holder so that the camera captured the entire truck and board.
Each participant spent a short time practicing with the guidance feature, with help provided by the experimenter as needed. (The experimenter chose practice hotspots that were different from the pre-defined K=10 hotspots.) In the practice phase, the experimenter explained that barriers sometimes block the user from moving in the suggested guidance direction. For instance, if in \ref{fig:exp} the system directs the participant to move “forward” towards the left side mirror, the stylus will soon hit the cabin roof/wall as it moves towards the mirror; the participant must find ways around such obstacles to find the target hotspot.

Next each participant performed the formal experimental trials, which consisted of the 2K=20 trial sequence described above. This sequence included 10 trials of both conditions (G and NG), for a total of 20 trials per participant. After completing the 20 search trials, we administered a System Usability Scale (SUS) \cite{brooke1996sus} to estimate the usability of the guidance system. Finally, we conducted a semi-structured interview asking what participants liked and disliked about the system, what needed to be improved, and how they might envision an ideal guidance system. Since P4 is a co-author on this paper, we did not administer the SUS or semi-structured interview to him.

# Experimental Results and Analysis {#results}

We conducted an experiment to assess the feasibility of the spatial guidance feedback and to determine whether the feedback speeds up the target search. The experiment (\ref{fig:exp}) compared how much time a blind participant needed to find hotspots (points of interest) on a 3D object with spatial guidance vs. without it. The object was a large toy fire truck (27 cm x 12 cm x 14 cm), with K=10 distinct hotspot locations defined in advance. The hotspots were tactilely salient features such as lights, hooks or other features that protruded from the truck. Some hotspots were in recessed locations on the truck, such as the shift stick and steering wheel (\ref{fig:exp}) in the front cabin, or on the side of the truck facing away from the participant; these hotspots were difficult to find and access, whether by fingertip or by stylus. In fact, the difficulty of accessing such locations highlights an advantage of the CamIO stylus, whose tip need not be visible to the camera. This allows the user to explore recessed areas or areas that are otherwise invisible to the camera.

\begin{figure}
\includegraphics[width=5.5cm, keepaspectratio,]{res/img/fig2a.jpg}
\includegraphics[width=6.5cm, keepaspectratio,]{res/img/fig2b.jpg}
\caption{(left) Plot of time (x axis) to find target hotspot location for G (‘o’ symbol) vs. NG  (‘x’ symbol) conditions, with time-out declared after 120 sec. of unsuccessful search. Results are shown for each participant (indicated on the y axis and also using color) and at all 10 target locations. (right) Histogram plot showing the distribution of NG time minus G time. The plot aggregates time differences for all participants over all locations; positive times mean the search was faster with G than with NG.}
\label{fig:res}
\end{figure}

We analyzed the differences between times under G vs. NG using a Wilcoxon signed rank test. The resulting p-value is p = 3.8 x 10-8. This is a highly statistically significant p-value supporting the rejection of the null hypothesis that the distribution of times under G is the same as the distribution of times under NG.

The following System Usability Scale (SUS) scores were calculated: P1=65, P2=75, P3=77.5, for an average of 72.5. An SUS score of 68 is considered “average” while scores in the mid- to high 70s are considered “good” (\url{https://measuringu.com/interpret-sus-score/}). Overall, the SUS scores suggest that the guidance system is usable but needs improvement.

Next we summarize the qualitative feedback we obtained in the semi-structured interviews. All participants who provided qualitative feedback (P1, P2 and P3) acknowledged the benefit of the audio guidance that was provided, which they found intuitive and clearly superior to finding hotspots without guidance. However, P1 and P2 felt that the limited spatial accuracy of the CamIO system was the biggest problem that needed to be addressed; for instance, in some cases the system announced the target had been reached even though the stylus tip was hovering about 1 cm above the truck surface. (This spatial inaccuracy results from noise in the board pose and stylus tip location estimates.) P3 expressed frustration with time lag, i.e., the voice feedback tended to lag behind the current location of the stylus, even when she was moving the stylus slowly; she also felt that the speed of the repetitive speech feedback made her feel rushed.

When asked what improvements the participants wanted, P1 suggested equipping the stylus with a button that turns on the guidance feedback, and the addition of multiple vibrators on the stylus that could provide haptic feedback indicating which way the stylus should be moved. P2 suggested adding audio information similar to the earlier guidance system that the authors experimented with (Sec. 3), which would use a combination of tempo, pitch and volume changes to signal when the stylus gets closer to the target location. P3 expressed the desire that CamIO work with any ordinary pen or pencil, rather than requiring the special stylus we used.

# Conclusion {#conclusion}

We have devised a 3D spatial guidance system that helps a blind user find a specific location of interest on an object using CamIO, an audio-based AR system. The guidance system uses simple verbal commands (“left”/ “right”/“up”/etc.) to tell the user which direction to move the stylus tip. Experiments with four blind participants show that the guidance system significantly speeds up the search for target hotspots compared with an exhaustive search. A secondary contribution of this study is to demonstrate the ability of the CamIO stylus to access recessed or otherwise obscured locations on an object that are not themselves visible to the camera.

While we applied this guidance approach specifically to CamIO, it could be useful in any AR system, whether the system tracks the user’s hands, handheld stylus or other pointing tool. Such audio-based guidance is not only helpful for blind users but is also useful in applications where visibility is limited (either by environmental conditions or because of a visual impairment), or for sighted users who prefer a cross-sensory display (e.g., audio and visual combined).

Future work will focus on refining and improving the audio interface. One possible way to speed up the guidance process is to add more specific directional feedback, e.g., “far left” would tell the user to move the stylus farther to the left than the “left” instruction. The verbal directions could be augmented with 3D spatialized sound as in \cite{mayAuditoryDisplaysFacilitate2019}, which could make the guidance more intuitive. Finally, we will continue to test and improve the CamIO system, with an emphasis on refining the spatial accuracy of the tip location estimates using improved computer vision algorithms.