### 2020-05-28 Manuscript.tex

% This is template.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04

\documentclass[runningheads]{res/templates/llncs}

\usepackage{graphicx}
\usepackage[utf8]{inputenc}
\usepackage[hyphens]{url}
\usepackage{hyperref}
\usepackage{etoolbox}

% Include packages from YAML header

\begin{document}

\title{An Audio-Based 3D Spatial Guidance AR System for Blind Users}

\titlerunning{Audio-Based 3D Spatial Guidance}
%
\newbool{flag}
\boolfalse{flag}
%
\author{
\ifbool{flag}{\and}{\booltrue{flag}}James Coughlan\inst{1}\orcidID{0000-1111-2222-3333}
\ifbool{flag}{\and}{\booltrue{flag}}Brandon Biggs\inst{1}\orcidID{0000-1111-2222-3333}
\ifbool{flag}{\and}{\booltrue{flag}}Marc-Aurèle Rivière\inst{2}\orcidID{0000-0002-5108-3382}
\ifbool{flag}{\and}{\booltrue{flag}}Huiying Shen\inst{1}\orcidID{0000-1111-2222-3333}
}
%
\authorrunning{J. Couhlan et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\boolfalse{flag}
%
\institute{
\ifbool{flag}{\and}{\booltrue{flag}}
 The Smith-Kettlewell Eye Research Institute, San Francisco, CA \\
 \email{\href{mailto:coughlan@ski.org}{\nolinkurl{coughlan@ski.org}},
\href{mailto:brandon.biggs@ski.org}{\nolinkurl{brandon.biggs@ski.org}},
\href{mailto:hshen@ski.org}{\nolinkurl{hshen@ski.org}}} 
\ifbool{flag}{\and}{\booltrue{flag}}
 LITIS, University of Rouen, Normandy, France \\
 \email{\href{mailto:marc-aurele.riviere@univ-rouen.fr}{\nolinkurl{marc-aurele.riviere@univ-rouen.fr}}} 
}

%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Augmented reality (AR) has great potential for blind users because it
enables a range of applications that provide audio information about
specific locations or directions in the user's environment. For
instance, the CamIO (``Camera Input-Output'') AR app makes physical
objects (such as documents, maps, devices and 3D models) accessible to
blind and visually impaired persons by providing real-time audio
feedback in response to the location on an object that the user is
touching (using an inexpensive stylus). An important feature needed by
blind users of AR apps such as CamIO is a 3D spatial guidance feature
that provides real-time audio feedback to help the user find a desired
location on an object. We have devised a simple audio interface to
provide verbal guidance towards a target of interest in 3D. The
experiment we report with blind participants using this guidance
interface demonstrates the feasibility of the approach and its benefit
for helping users find locations of interest.
\\
\boolfalse{flag}
\keywords{
\ifbool{flag}{\and}{\booltrue{flag}}
Assisitive Devices
\ifbool{flag}{\and}{\booltrue{flag}}
Augmented Reality
\ifbool{flag}{\and}{\booltrue{flag}}
Auditory Substitution
\ifbool{flag}{\and}{\booltrue{flag}}
Visual Impairment
}
\end{abstract}
%
%
\hypertarget{sota}{%
\section{State of the Art and Related Technology}\label{sota}}

Many people who are blind or visually impaired have difficulties
accessing a range of everyday objects, including printed documents,
maps, infographics, appliances and 3D models used in STEM education,
needed for daily activities in schools, the home and the workplace. This
limits their participation in many cultural, professional and
educational activities. While braille labeling is often a useful means
of providing access to such objects, there is only limited space for
braille, and this method is only accessible to those who can read
braille.

Audio labels are a powerful supplement or alternative to braille labels.
Standard methods of implementing audio labels require the use of special
hardware and materials, which is costly and limits their adoption. For
instance, tactile graphics may be overlaid on a touch-sensitive tablet
\cite{TalkingTactileTablet}. Documents or other surfaces can be covered
with a special form of paper readable by a ``smart pen''
\cite{mieleTalkingTactileApps2010}. The PenFriend
\cite{kendrickPenFriendTouchMemo2011} stylus uses small barcodes affixed
to an object to define audio labels.

There is growing interest in using computer vision-based approaches for
audio labeling. In these approaches a camera tracks the user's fingers
or handheld pointer as they explore an object, enabling audio labeling
for existing objects with minimal or no customization. Past computer
vision approaches to audio labeling include
\cite{shiMarkitTalkitLowBarrier2017}, which focuses on 3D printed
objects, \cite{thevinAugmentedRealityPeople2018}, which uses a depth
camera to access flat documents, and
\cite{fuscoTactileGraphicsHelper2015}, which provides access to tactile
graphics to students in educational settings.

We build on our past work on CamIO
\cite{shenCamIO3DComputer2013,coughlanEvaluatingAuthorUser2017,coughlanTowardsAccessibleAudioLabeling2020},
which is short for ``Camera Input-Output.'' We have implemented CamIO as
a stand-alone iPhone app that tracks an inexpensive wooden stylus held
by the user to point at locations of interest on an object (called
hotspots). When the user points at a hotspot with the stylus tip,
relevant audio information, such as text-to-speech or other sounds, is
announced.

Experiments with blind participants have demonstrated the feasibility of
using CamIO to access existing audio labels. Moreover, our approach
allows a visually impaired user to create audio labels independently
\cite{coughlanTowardsAccessibleAudioLabeling2020}. However, an ability
that CamIO still lacks is the ability to provide real-time guidance to a
hotspot. Without such guidance, the user must search the entire surface
of an object with the stylus tip until they hear the hotspot they are
seeking. This need is particularly acute in cases where the surface is
lacking in tactile cues, such as the touch screen buttons on a microwave
oven \cite{coughlanTowardsAccessibleAudioLabeling2020}.

Guidance not only saves the user time in exploring large or intricate
objects, or objects lacking in tactile features, but can enable entirely
new ways of interacting with objects. For instance, a geography learning
app could ask a student to point to a specific city on a world globe,
and if the student has trouble finding it, the app could guide them to
it in an intuitive way. Other examples could include guiding a user to a
specific button on an appliance (such as a microwave) to perform a
desired function, guiding a student to a specific element on the
periodic table, or even guiding a repairperson to find a part of a
machine with limited visibility. Our guidance system is inspired by
recent work on auditory and haptic displays for 3D spatial guidance
\cite{mayAuditoryDisplaysFacilitate2019,guezou-philippePrototypingEvaluatingSensory2018},
which provide real-time audio/haptic feedback to guide a user towards a
target in a virtual environment. We initially developed an interface
similar to \cite{mayAuditoryDisplaysFacilitate2019} for CamIO, but after
preliminary testing we devised a simpler and more effective interface
for our application, which we evaluated experimentally with four blind
participants.

\hypertarget{overview}{%
\section{Overview of the CamIO System}\label{overview}}

CamIO is an augmented reality (AR) iPhone app that uses computer vision
to provide audio labels for rigid 3D objects. The rigid object of
interest is mounted on a board, which is a printed barcode pattern
(\ref{fig:exp}) that allows CamIO to estimate the object's pose (3D
translation and 3D orientation) relative to the camera. The iPhone is
typically mounted on a tripod or a gooseneck cellphone holder
\ref{fig:exp} to obtain a clear view of the board and object, but in
some cases \cite{coughlanTowardsAccessibleAudioLabeling2020} users may
hold the iPhone by hand. The user points the tip of a passive wooden
stylus \ref{fig:exp}, covered with another barcode pattern, to different
locations on an object. CamIO estimates the stylus tip's 3D location
relative to the camera and uses the pose estimate of the board to
determine the stylus tip's 3D location relative to the object. Whenever
the tip is close enough to a pre-defined hotspot, it triggers audio
information about the hotspot. (In other words, hotspot locations are
defined relative to the board, which is mounted rigidly to the object.)

A second stylus (not shown, identical in form but distinct from the one
shown in \ref{fig:exp}) is used to create new hotspots, and allows the
user to make an audio recording to associate with each new hotspot. Our
recent work \cite{coughlanTowardsAccessibleAudioLabeling2020}
demonstrates the accessibility of this authoring process, which allows
blind users to create their own annotations (hotspots and associated
audio labels) independently. However, in the study presented in this
paper, the experimenter was the only one who created hotspots, since the
emphasis here is on providing guidance to pre-existing hotspots.

\hypertarget{material}{%
\section{3D Spatial Guidance Interface}\label{material}}

We build on the auditory display described in
\cite{mayAuditoryDisplaysFacilitate2019}, which is intended for
audio-based 3D targeting of locations in a large virtual workspace (such
as a kitchen). Accordingly, the first incarnation of our guidance
approach used a repeating monaural sound, whose tempo indicated how far
the stylus tip is from the target, and with a pitch (low, medium or
high) that indicated whether the tip needed to move up or down (or else
was aligned properly in the vertical direction).

Informal tests with two blind participants (one female of age 42, who is
also participant P1 in the experiment described in the next section, and
one male of age 73) showed that this guidance worked satisfactorily in
some cases. However, we found that the feedback was too indirect: the
participants didn't always know how to move the stylus to increase the
tempo. As a result, the resulting guidance process was unacceptably
slow. Thus we devised a second, more direct guidance approach: verbal
directions telling the user to move the stylus along the cardinal
directions (up, down, left, right, forward or back). While the optimal
direction to move the stylus tip towards the hotspot might entail a
combination of these directions (e.g., simultaneously move left and up),
for simplicity we issue directions along only one cardinal direction at
a time -- the direction that will most improve the alignment to the
target.

We define the guidance interface as follows. The board (barcode pattern
beneath the truck, see \ref{fig:exp}) defines an xyz coordinate frame,
with \(+x\) = right, \(-x\) = left, \(+y\) = forward (away from the
user), \(-y\) = back (towards the user), \(+z\) = up and \(-z\) = down.
Let the desired target hotspot location be denoted \((x*, y*, z*)\), and
the current location of the stylus tip be \((x,y,z)\). Define the error
vector \(e = (x-x*, y-y*, z-z*)\), and let index \(a = argmax_i |e_i|\),
i.e., a is the index (1, 2 or 3, corresponding to x, y or z,
respectively) of the entry in e with the highest absolute value. In
other words, a indicates which coordinate (x, y or z) is most discrepant
from the target coordinate location \((x*, y*, z*)\). Then the
appropriate direction is issued, e.g., ``left'' if \(a = 1\) and
\(x > x*\), ``right'' if \(a = 1\) and \(x < x*\), ``up'' if \(a = 3\)
and \(z < z*\), etc. Directions are announced roughly twice per second
whenever the stylus is visible. The target hotspot is announced when the
stylus gets to within approximately 1 cm from the hotspot. No audio
feedback is issued when the stylus is not visible.

\hypertarget{exp}{%
\section{Experiment with Blind Participants}\label{exp}}

We conducted an experiment to assess the feasibility of the spatial
guidance feedback and to determine whether the feedback speeds up the
target search. The experiment \ref{fig:exp} compared how much time a
blind participant needed to find hotspots (points of interest) on a 3D
object with spatial guidance vs.~without it. The object was a large toy
fire truck (27 cm x 12 cm x 14 cm), with K=10 distinct hotspot locations
defined in advance. The hotspots were tactilely salient features such as
lights, hooks or other features that protruded from the truck. Some
hotspots were in recessed locations on the truck, such as the shift
stick and steering wheel (\ref{fig:exp}) in the front cabin, or on the
side of the truck facing away from the participant; these hotspots were
difficult to find and access, whether by fingertip or by stylus. In
fact, the difficulty of accessing such locations highlights an advantage
of the CamIO stylus, whose tip need not be visible to the camera. This
allows the user to explore recessed areas or areas that are otherwise
invisible to the camera.

\begin{figure}
\includegraphics[width=0.55\linewidth]{res/img/fig1a} \includegraphics[width=0.45\linewidth]{res/img/fig1b} \caption{(left) Blind participant uses CamIO to explore a toy fire truck. The participant holds the CamIO stylus (a barcode pattern printed on a square wooden dowel with a pointy tip) to point to possible hotspots on the truck. The truck is glued to the board, a piece of paper with a separate barcode pattern, which is taped to the table. An iPhone 8 running CamIO is held by a gooseneck cellphone holder to capture a view of the truck, board and stylus. (right) Stylus points to steering wheel in the truck cabin. Though the steering wheel is in a recessed location that is not visible to CamIO, the stylus tip location can still be estimated.}\label{fig:exp}
\end{figure}

Each hotspot location was searched for under two feedback conditions: G
(guidance provided) and NG (no guidance). The participant was given up
to 120 sec.~to find each hotspot, after which a time-out was declared.
In order to minimize the influence of any possible learning effects,
which could make it easier to find a hotspot the second time it was
searched for, we randomized the sequence of hotspots and alternated
between NG and G trials. This resulted in a sequence of 2K = 20 trials,
starting with an NG trial, the same sequence for each participant.

We had a total of N=4 blind participants in our experiment (2 male/2
female, ages from 26-42). Participants P1, P3 and P4 were already
familiar with the CamIO system (P4 is a co-author of this paper) while
P2 was unfamiliar with it. None of the participants was told the study
design before their trials, and we didn't tell them whether or not any
locations would be repeated in the 20 trials. After obtaining IRB
consent, we reviewed how CamIO works if necessary and demonstrated the
audio guidance feature. The experimenter launched CamIO on an iPhone 8
and positioned it on a gooseneck cellphone holder so that the camera
captured the entire truck and board. Each participant spent a short time
practicing with the guidance feature, with help provided by the
experimenter as needed. (The experimenter chose practice hotspots that
were different from the pre-defined K=10 hotspots.) In the practice
phase, the experimenter explained that barriers sometimes block the user
from moving in the suggested guidance direction. For instance, if in
\ref{fig:exp} the system directs the participant to move ``forward''
towards the left side mirror, the stylus will soon hit the cabin
roof/wall as it moves towards the mirror; the participant must find ways
around such obstacles to find the target hotspot.

Next each participant performed the formal experimental trials, which
consisted of the 2K=20 trial sequence described above. This sequence
included 10 trials of both conditions (G and NG), for a total of 20
trials per participant. After completing the 20 search trials, we
administered a System Usability Scale (SUS) \cite{brooke1996sus} to
estimate the usability of the guidance system. Finally, we conducted a
semi-structured interview asking what participants liked and disliked
about the system, what needed to be improved, and how they might
envision an ideal guidance system. Since P4 is a co-author on this
paper, we did not administer the SUS or semi-structured interview to
him.

\hypertarget{results}{%
\section{Experimental Results and Analysis}\label{results}}

We conducted an experiment to assess the feasibility of the spatial
guidance feedback and to determine whether the feedback speeds up the
target search. The experiment (\ref{fig:exp}) compared how much time a
blind participant needed to find hotspots (points of interest) on a 3D
object with spatial guidance vs.~without it. The object was a large toy
fire truck (27 cm x 12 cm x 14 cm), with K=10 distinct hotspot locations
defined in advance. The hotspots were tactilely salient features such as
lights, hooks or other features that protruded from the truck. Some
hotspots were in recessed locations on the truck, such as the shift
stick and steering wheel (\ref{fig:exp}) in the front cabin, or on the
side of the truck facing away from the participant; these hotspots were
difficult to find and access, whether by fingertip or by stylus. In
fact, the difficulty of accessing such locations highlights an advantage
of the CamIO stylus, whose tip need not be visible to the camera. This
allows the user to explore recessed areas or areas that are otherwise
invisible to the camera.

\begin{figure}
\includegraphics[width=5.5cm, keepaspectratio,]{res/img/fig2a.jpg}
\includegraphics[width=6.5cm, keepaspectratio,]{res/img/fig2b.jpg}
\caption{(left) Plot of time (x axis) to find target hotspot location for G (‘o’ symbol) vs. NG  (‘x’ symbol) conditions, with time-out declared after 120 sec. of unsuccessful search. Results are shown for each participant (indicated on the y axis and also using color) and at all 10 target locations. (right) Histogram plot showing the distribution of NG time minus G time. The plot aggregates time differences for all participants over all locations; positive times mean the search was faster with G than with NG.}
\label{fig:res}
\end{figure}

We analyzed the differences between times under G vs.~NG using a
Wilcoxon signed rank test. The resulting p-value is p = 3.8 x 10-8. This
is a highly statistically significant p-value supporting the rejection
of the null hypothesis that the distribution of times under G is the
same as the distribution of times under NG.

The following System Usability Scale (SUS) scores were calculated:
P1=65, P2=75, P3=77.5, for an average of 72.5. An SUS score of 68 is
considered ``average'' while scores in the mid- to high 70s are
considered ``good'' (\url{https://measuringu.com/interpret-sus-score/}).
Overall, the SUS scores suggest that the guidance system is usable but
needs improvement.

Next we summarize the qualitative feedback we obtained in the
semi-structured interviews. All participants who provided qualitative
feedback (P1, P2 and P3) acknowledged the benefit of the audio guidance
that was provided, which they found intuitive and clearly superior to
finding hotspots without guidance. However, P1 and P2 felt that the
limited spatial accuracy of the CamIO system was the biggest problem
that needed to be addressed; for instance, in some cases the system
announced the target had been reached even though the stylus tip was
hovering about 1 cm above the truck surface. (This spatial inaccuracy
results from noise in the board pose and stylus tip location estimates.)
P3 expressed frustration with time lag, i.e., the voice feedback tended
to lag behind the current location of the stylus, even when she was
moving the stylus slowly; she also felt that the speed of the repetitive
speech feedback made her feel rushed.

When asked what improvements the participants wanted, P1 suggested
equipping the stylus with a button that turns on the guidance feedback,
and the addition of multiple vibrators on the stylus that could provide
haptic feedback indicating which way the stylus should be moved. P2
suggested adding audio information similar to the earlier guidance
system that the authors experimented with (Sec. 3), which would use a
combination of tempo, pitch and volume changes to signal when the stylus
gets closer to the target location. P3 expressed the desire that CamIO
work with any ordinary pen or pencil, rather than requiring the special
stylus we used.

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

We have devised a 3D spatial guidance system that helps a blind user
find a specific location of interest on an object using CamIO, an
audio-based AR system. The guidance system uses simple verbal commands
(``left''/ ``right''/``up''/etc.) to tell the user which direction to
move the stylus tip. Experiments with four blind participants show that
the guidance system significantly speeds up the search for target
hotspots compared with an exhaustive search. A secondary contribution of
this study is to demonstrate the ability of the CamIO stylus to access
recessed or otherwise obscured locations on an object that are not
themselves visible to the camera.

While we applied this guidance approach specifically to CamIO, it could
be useful in any AR system, whether the system tracks the user's hands,
handheld stylus or other pointing tool. Such audio-based guidance is not
only helpful for blind users but is also useful in applications where
visibility is limited (either by environmental conditions or because of
a visual impairment), or for sighted users who prefer a cross-sensory
display (e.g., audio and visual combined).

Future work will focus on refining and improving the audio interface.
One possible way to speed up the guidance process is to add more
specific directional feedback, e.g., ``far left'' would tell the user to
move the stylus farther to the left than the ``left'' instruction. The
verbal directions could be augmented with 3D spatialized sound as in
\cite{mayAuditoryDisplaysFacilitate2019}, which could make the guidance
more intuitive. Finally, we will continue to test and improve the CamIO
system, with an emphasis on refining the spatial accuracy of the tip
location estimates using improved computer vision algorithms.
%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
\bibliographystyle{res/templates/splncs04}
% \bibliography{bibliography}
\bibliography{res/refs.bib}

\end{document}