### 2020-05-29 R/Index.Rmd

---
title: "CamIO data analysis"
author: "RIVIERE Marc-Aurèle"
output:
  prettydoc::html_pretty:
    highlight: github
    math: mathjax
    code_folding: hide
    number_sections: yes
    theme: cayman
    toc: yes
  html_notebook:
    code_folding: hide
    highlight: tango
    number_sections: yes
    theme: spacelab
    toc: yes
  pdf_document:
    number_sections: yes
    toc: yes
  html_document:
    theme: sandstone
    highlight: tango
    df_print: paged
    toc: yes
always_allow_html: yes
---

# Setup:

```{r setup, include=FALSE, cache=FALSE}

if (!require(pacman)) install.packages("pacman")

pkg_list <- c("tidyverse", "glue", "broom", "magrittr", "prettydoc", "DT", "reshape2", "plotly", "moments", "emmeans", "gtsummary", "lmtest", "table1", "rstatix", "lme4", "ggeffects", "conflicted", "forecast")

pacman::p_load(char = pkg_list) #update=TRUE
pacman::p_load_gh(c("rstudio/gt", "tidyverse/dplyr"))
#knitr::write_bib(c(.packages(), pkg_list), 'bib/packages.bib')

knitr::opts_chunk$set(
  fig.align="center",
  #fig.pos = "!H",
  #out.extra = "",
  #tidy = TRUE,
  warning = FALSE,
  echo = FALSE,
  message = FALSE,
  error=FALSE,
  cache = FALSE
)

options(scipen=999, digits=5)

tc <- 0
alpha <- 0.05
set.seed(42)

initial.data <- "../data/initial.csv"
```

```{r}
sessionInfo()

conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
conflict_prefer("layout", "plotly")

# includes:
#   in_header: header.html
#   after_body: sig.html
```

# Data preparation:
***
***

```{r}
data.wide <- readr::read_csv(initial.data) %>% 
  dplyr::select(-X1) %>% 
  mutate(across(c(hotspot, mode), as.factor)) %T>% print
```

```{r results=FALSE}
### With reshape2
data.long <- data.wide %>% 
  reshape2::melt(measure.vars=c("P1", "P2", "P3", "P4"), variable.name = "subject", value.name="time") %>% 
  group_by(subject) %>% mutate(id = seq_along(hotspot), .before = hotspot) %>% ungroup() %T>% print

### (old) With tidyr 0.8.5
#data.long <- data.wide %>% gather(key="subject", value="time", "P1":"P4") %T>% print
#data.long <- data.wide %>% gather(key="subject", value="time", c("P1", "P2", "P3", "P4")) %T>% print

### With tidyverse >= 1.0.0 && dplyr >= 0.8.99
# data.long <- data.wide %>% 
#   mutate(id = seq_along(hotspot), .before = hotspot) %>% 
#   pivot_longer(cols = "P1":"P4", names_to="subject", names_transform = list(subject = as.factor), values_to = "time") %>%
#   group_by(subject) %>% arrange(id, .by_group=T) %>% ungroup() %T>% print
```

# Summary / descriptive statistics:
***
***

```{r child = '01-visualizations.Rmd', echo=TRUE}
```

# Analysis:
***
***

```{r child = '02-analysis.Rmd', echo=TRUE}
```