### 2020-06-09 R/01-visualizations.Rmd
Update figures

## Data:
***

Original data:

```{r}
DT::datatable(
  data.wide,
  class = 'cell-border stripe compact',
  options = list(pageLength = 20, autoWidth = TRUE),
  caption = htmltools::tags$caption(
    style = 'caption-side: bottom; text-align: center;',
    glue('Fig {tc<-tc+1}: Filtered data (wide format)'))
)
```

Long format:

```{r}
DT::datatable(
  data.long,
  class = 'cell-border stripe compact',
  filter = 'top',
  options = list(pageLength = 20, autoWidth = TRUE),
  caption = htmltools::tags$caption(
    style = 'caption-side: bottom; text-align: center;',
    glue('Fig {tc<-tc+1}: Data (long format)'))
)
```

## Global summary statistics:
***

```{r warnings=F}
library(pillar)
options(pillar.sigfig=3)

desc <- data.long %>%
  group_by(mode) %>%
  summarize(
    n = n(),
    median = median(time),
    mean = mean(time),
    sd = sd(time),
    skew = skewness(time),
    kurt = kurtosis(time),
    SW.p = shapiro.test(time)$p.value
  ) %>% mutate(across(where(is.character), as.numeric))

td <- tabl(desc)

DT::datatable(
  desc,
  class = 'cell-border stripe compact',
  caption = htmltools::tags$caption(
    style = 'caption-side: bottom; text-align: center;',
    glue('Fig {tc<-tc+1}: Summary statistics'))
) %>% 
  formatRound(4:ncol(desc)-1, 2) %>%
  formatRound(ncol(desc), 4)
  # formatStyle(
  #   'SW.p',
  #   color = styleInterval(alpha, c('red', 'gray')),
  #   fontWeight = styleInterval(alpha, c('bold', 'normal')),
  # )

# png("../res/img/tab1.png", width=480, height=480, bg = "white")
# grid.table(desc)
# dev.off()
```

[!] TODO: better summary table
|   See [here](https://dabblingwithdata.wordpress.com/2018/01/02/my-favourite-r-package-for-summarising-data/)

```{r}
#data.long %>% tbl_summary %>% bold_labels
#summary(arsenal::tableby(mode ~ hotspot, data.long), title = "Arsenal summary")
table1::table1(~time | mode + subject, data = data.long)
```

## Summary statistics per guidance mode:
***

### Distribution of scores per guidance mode :

```{r}
fig <- data.long %>% 
  group_by(mode) %>% 
  mutate(med = median(time)) %>% 
  ungroup() %>%
  ggplot(aes(x=time, fill=mode, color=mode)) +
  geom_histogram(binwidth=12, fill="white", boundary = 1) + #aes(y=..density..)
  geom_density(alpha=0.6, adjust=6) + 
  theme(legend.position = "none") +
  labs(x = "Time to find hotspot", y="Count") +
  facet_wrap(. ~ mode) +
  geom_vline(aes(xintercept=med, group=mode, color=mode), linetype="dashed", size=1) +
  geom_text(aes(x=med - 10, label=med, y=-1.5))

ggsave(plot=fig, filename="../res/img/histNG.png", device="png", dpi=320, scale=1)
ggplotly(fig)
```

### Box-Plots per subject & guidance mode:

```{r}
plot_ly(
    data.long,
    x = ~subject:mode,
    y = ~time,
    color = ~subject:mode,
    type = 'box',
    showlegend = FALSE,
    colors = "Set1"
  ) %>%
  plotly::layout(
    xaxis = list(title = "Subjects"),
    yaxis = list(title = "Time", range = c(0,120)),
    title = glue('Fig {tc<-tc+1}: Violin + Box-Plots per guidance mode')
  )
```

### Lineplot of averages between guidance mode :

```{r}
dat <- data.long %>% Rmisc::summarySE(measurevar="time", groupvars=c("mode", "subject")) %>% print
#dat2 <- data.long %>% mutate(se = se(time))
pd <- position_dodge(0.15)

fig <- dat %>%
  ggplot(aes(x = mode, y = time, color = subject)) + #group=1
  geom_line(position = pd) +
  geom_errorbar(aes(ymin = time - se, ymax = time + se), width = .1, position = pd) +
  geom_point(shape=21, size=2, fill="white", position = pd) +
  labs(x="Guidance mode", y="Time to find hotspot")

ggplotly(fig)
ggsave(plot=fig, filename="../res/img/lines.png", device="png", dpi=320, scale=1)
```

### Summary of scores per guidance mode :

```{r}
summary.mode <- data.long %>%
  group_by(mode) %>%
  summarize(
    n = n(),
    median = median(time),
    mean = mean(time),
    skew = skewness(time),
    kurt = kurtosis(time),
    SW.p = shapiro.test(time)$p.value
  )

DT::datatable(
  summary.mode,
  class = 'cell-border stripe compact',
  caption = htmltools::tags$caption(style = 'caption-side: bottom; text-align: center;',
                                    glue('Fig {tc<-tc+1}: Summary statistics per guidance mode'))
) %>%
  formatRound(5:ncol(summary.mode), 4) %>%
  formatStyle('SW.p',
              color = styleInterval(alpha, c('red', 'gray')),
              fontWeight = styleInterval(alpha, c('bold', 'normal')))
```

## Summary statistics per subject:
***

### Distribution per subject (guided condition):

```{r}
fig <- data.long %>% 
  filter(mode == "G") %>%
  group_by(subject) %>% 
  mutate(med = median(time)) %>% 
  ungroup() %>%
  ggplot(aes(x=time, fill=subject, color=subject)) +
  geom_histogram(aes(y=..density..), binwidth=5, alpha=0.8, fill="white", boundary = 1, position="identity") +
  geom_density(aes(fill=subject), alpha=0.6, adjust=2) + 
  theme(legend.position = "none") +
  facet_wrap(. ~ subject, ncol=2) +
  geom_vline(aes(xintercept=med, color=subject), linetype="dashed", size=1) +
  geom_text(aes(x=med - 5, label=med, y=-0.01)) +
  labs(title= glue('Fig {tc<-tc+1}: Score distributions & median scores per subject (guided mode)'), x = "Time to find hotspot")

ggplotly(fig)
```

### Violin & Box plots per subject (guided condition):

```{r}
plot_ly(
    data.long %>% filter(mode == "G"),
    x = ~subject,
    y = ~time,
    color = ~subject,
    type = 'violin',
    points = 'all',
    box = list(visible = T),
    meanline = list(visible = T),
    showlegend = FALSE,
    colors = "Set1"
  ) %>%
  plotly::layout(
    xaxis = list(title = "Subjects"),
    yaxis = list(title = "Time", range = c(0,120)),
    title = glue('Fig {tc<-tc+1}: Violin + Box-Plots (guided condition)')
  )
```

### Summary per subject (guided condition):

```{r}
summary.subject <- data.long %>%
  filter(mode == "G") %>%
  group_by(subject) %>%
  summarize(
    n = n(),
    median = median(time),
    mean = mean(time),
    skew = skewness(time),
    kurt = kurtosis(time),
    SW.p = shapiro.test(time)$p.value
  )

DT::datatable(
  summary.subject,
  class = 'cell-border stripe compact',
  caption = htmltools::tags$caption(style = 'caption-side: bottom; text-align: center;',
                                    glue('Fig {tc<-tc+1}: Summary statistics per subject'))
) %>%
  formatRound(5:ncol(summary.subject), 4) %>%
  formatStyle('SW.p',
              color = styleInterval(alpha, c('red', 'gray')),
              fontWeight = styleInterval(alpha, c('bold', 'normal')))
```

```{r}
## Violin + Box plots per subject * mode

# violin <- data.long %>%
#   plot_ly(
#     type = 'violin',
#     points = 'all',
#     #color = ~subject,
#     box = list(visible = T),
#     meanline = list(visible = T),
#     showlegend = FALSE,
#     colors = "Set1"
#   )
# 
# violin %<>%
#   add_trace(
#     x = ~subject[data.long$mode == 'G'],
#     y = ~time[data.long$mode == 'G'],
#     legendgroup = 'G',
#     scalegroup = 'G',
#     name = 'G',
#     color = I("blue")
#     # box = list(
#     #   visible = T
#     # ),
#     # meanline = list(
#     #   visible = T
#     # )
#   )
# 
# violin %<>%
#   add_trace(
#     x = ~subject[data.long$mode == 'N'],
#     y = ~time[data.long$mode == 'N'],
#     legendgroup = 'N',
#     scalegroup = 'N',
#     name = 'N',
#     color = I("pink")
#     # box = list(
#     #   visible = T
#     # ),
#     # meanline = list(
#     #   visible = T
#     # ),
#   ) 
# 
# violin %<>%
#   layout(
#     xaxis = list(title = "Subject and Guidance mode"),
#     yaxis = list(title = "Time", range = c(0,130)),
#     title = glue('Figure {tc<-tc+1}: Violin + Box-Plots for both conditions'),
#     violinmode = 'group'
#   )
# 
# violin
```

## Summary statistics per hotspot:
***

### Distribution per hotspot (guided condition):

```{r fig.width=10, fig.height=10}
fig <- data.long %>%
  filter(mode == "G") %>%
  group_by(hotspot) %>% 
  mutate(med = median(time)) %>% 
  ungroup() %>%
  ggplot(aes(x=time, fill=hotspot, color=hotspot)) +
  geom_histogram(aes(y=..density..), binwidth=5, alpha=0.8, fill="white", boundary = 1, position="identity") +
  geom_density(aes(fill=hotspot), alpha=0.6, adjust=3) + 
  theme(legend.position = "none") +
  facet_wrap(. ~ hotspot, ncol=4) +
  geom_vline(aes(xintercept=med, color=hotspot), linetype="dashed", size=1) +
  geom_text(aes(x=med + 10, label=med, y=-0.02)) +
  labs(title = glue('Fig {tc<-tc+1}: Score distributions & median scores per hotspot (guided mode)'),
    x = "Time to find hotspot")

ggplotly(fig)
```

### Violin & Box plots per hotspot (guided mode):

```{r}
plot_ly(
    data.long %>% filter(mode == "G"),
    x = ~hotspot,
    y = ~time,
    color = ~hotspot,
    type = 'violin',
    points = 'all',
    box = list(visible = T),
    meanline = list(visible = T),
    showlegend = FALSE,
    colors = "Set1"
  ) %>%
  layout(
    xaxis = list(title = "Hotspots"),
    yaxis = list(title = "Time", range = c(0,120)),
    title = glue('Fig {tc<-tc+1}: Violin + Box-Plots per hotspot (guided mode)')
  )
```

### Summary per hotspot (guided mode):

```{r}
summary.hotspots <- data.long %>%
  filter(mode == "G") %>%
  group_by(hotspot) %>%
  summarize(
    n = n(),
    median = median(time),
    mean = mean(time),
    skew = skewness(time),
    kurt = kurtosis(time),
    SW.p = shapiro.test(time)$p.value
  )

DT::datatable(
  summary.hotspots,
  class = 'cell-border stripe compact',
  caption = htmltools::tags$caption(style = 'caption-side: bottom; text-align: center;',
                                    glue('Fig {tc<-tc+1}: Summary statistics per hotspot (guided mode)'))
) %>%
  formatRound(5:ncol(summary.hotspots), 4) %>%
  formatStyle('SW.p',
              color = styleInterval(alpha, c('red', 'black')),
              fontWeight = styleInterval(alpha, c('bold', 'normal')))
```