### 2020-06-10 R/ICCHP20.Rmd
Update analysis & figs

---
title: "R Notebook"
output: html_notebook
---

```{r include=FALSE}

if (!require(knitr)) install.packages("knitr")
if (!require(pacman)) install.packages("pacman")
pkg_list <- c("tidyverse", "glue", "magrittr", "plotly", "moments", "emmeans", "conflicted", "gridExtra", "kableExtra", "rcompanion", "broom", "rstatix", "ggpubr", "Rmisc")

pacman::p_load(char = pkg_list) #update=TRUE

knitr::opts_chunk$set(
  warning = FALSE,
  echo = FALSE,
  message = FALSE,
  error=FALSE,
  cache = FALSE
)

options(scipen=999, digits=4)

alpha <- 0.05
set.seed(42)

initial.data <- "../data/initial.csv"

conflict_prefer("mutate", "dplyr")
conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
conflict_prefer("summarize", "dplyr")
conflict_prefer("layout", "plotly")

tabl <- function(data, title="Table caption") {
  kable(data, col.names=gsub("[.]", " ", names(data)), booktabs=T, align="c", digits=3, format.args=list(big.mark=",", scientific=FALSE), escape=F, caption=title, linesep=c("")) %>% 
    kable_styling(font_size = 10, latex_options = c("striped", "scale_down", "hold_position"), position = "left") %>%
    row_spec(0, background = "#d4dbde", bold = T, color = "#2b4894")
}
options(knitr.table.format = ifelse(knitr::is_latex_output(), "latex", "pandoc"), knitr.kable.NA = "")
```

```{r results='hide'}
data.wide <- read_csv(initial.data) %>% 
  select(-X1) %>% 
  mutate(mode = ifelse(mode == "N", "NG", "G")) %>%
  mutate(across(c(hotspot, mode), as.factor))

data.long <- data.wide %>%
  mutate(id = seq_along(hotspot), .before = hotspot) %>%
  pivot_longer(cols = "P1":"P4", names_to="subject", names_transform = list(subject = as.factor), values_to = "time") %>%
  relocate(subject, id, mode, hotspot, time) %>%
  arrange(subject, id) %T>% print
```
***
# Wilcoxon
***

```{r}
data.long.wil <- data.long %>% arrange(subject, hotspot, mode) %T>% print
```

```{r}
wsrt1 <- wilcox.test(formula=time ~ mode, data=data.long.wil, paired=TRUE, alternative="two.sided")
wsrt1 %>% tidy
```

```{r}
G <- data.long.wil %>% filter(mode == "G") %>% select(time) %T>% print
NG <- data.long.wil %>% filter(mode == "NG") %>% select(time) %T>% print

wsrt2 <- wilcox.test(NG$time, G$time, paired=TRUE, alternative="two.sided")
wsrt2 %>% tidy
```

```{r}
wilcox_test(time ~ mode, data=data.long.wil, paired=TRUE, alternative="less")
```

```{r}
wrst.z <- wilcoxonZ(NG$time, G$time, paired = TRUE, digits = 3)
wrst.z
```

```{r}
wrst.z / sqrt(nrow(NG))
data.long.wil %>% rstatix::wilcox_effsize(formula=time ~ mode, paired=TRUE)
```
```{r}
sign_test(time ~ mode, data=data.long.wil)
```

# Success:

```{r}
data.long.logit <- data.long %>% mutate(success = ifelse(time >= 120, 0, 1)) %T>% print
```

```{r}
tbl1 <- table(data.long.logit$mode, data.long.logit$success)
tbl1
```

```{r}
tbl2 <- prop.table(tbl1) %>% round(2)
tbl2
```

```{r}
chisq_test(tbl1)
```
***
# Graphics
***

```{r}
hist1 <- data.long %>% gghistogram(x = "time", add = "median", rug = TRUE, color = "mode", fill = "mode", 
                          palette = c("#00AFBB", "#E7B800"), bins = 20, xlab="Time to find hotspot", ylab="Count")

hist2 <- data.long %>% gghistogram(x = "time", add = "median", rug = TRUE, color = "mode", fill = "mode", 
                          palette = c("#00AFBB", "#E7B800"), bins = 20, xlab="Time to find hotspot", ylab="Count", facet.by="mode")

ggsave("../res/img/hist_diff.png", hist1, device = "png", dpi=320)
ggsave("../res/img/hist_diff.png", hist2, device = "png", dpi=320)
```

```{r warnings=F}
data.pairwise <- data.long %>% group_by(subject, hotspot) %>%
  summarize(diff = abs(diff(time))) %>%
  ungroup() %T>% print
```

```{r}
data.pairwise %>% gghistogram(x = "diff", add = "median", rug = TRUE, bins = 20, 
                                           color="#FC4E07", fill="#FC4E07", xlab="Paired absolute time-differences", ylab="Count")

ggsave("../res/img/hist_diff.png", device = "png", dpi=320)
```

```{r}
data.long %>% arrange(subject, hotspot) %>% ggdotchart(
  x = "hotspot",
  y = "time",
  color = "mode",
  palette = c("#00AFBB", "#E7B800"),
  sorting = "ascending",
  add = "segments",
  add.params = list(color = "lightgray", size = 1.5),
  group = "mode",
  dot.size = 7,
  label = "time",
  font.label = list(
    color = "white",
    size = 8,
    vjust = 0.5
  ),
  facet.by = "subject",
  xlab="Hotspot",
  ylab="Time to find hotspot",
  ylim = c(0, 130)
)

ggsave("../res/img/full.png", device = "png", dpi=320)
```

```{r}
dat <- data.long %>% Rmisc::summarySE(measurevar="time", groupvars=c("mode", "subject")) %>% print
pd <- position_dodge(0.3)

conflict_prefer("arrange", "plyr")

fig <- dat %>%
  ggplot(aes(x = mode, y = time, color = subject)) + #group=1
  geom_line(position = pd) +
  geom_errorbar(aes(ymin = time - se, ymax = time + se), size = 0.9, width = 0.15, position = pd) +
  geom_point(aes(fill=subject), shape=21, size=3.2, position = pd) +
  theme_classic() +
  labs(x="Guidance mode", y="Time to find hotspot")

fig
ggsave("../res/img/lines.png", device = "png", dpi=320)
#ggplotly(fig)
```